\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-04-22 analysis/paper/paper.Rmd
last edits for draft
Fixed code formatting, typos, plot colors.

In the first two case studies, we described the basic functionalities of the \pkg{casebase} package: creating population-time plots, fitting parametric models for hazard functions, and estimating the corresponding cumulative incidence curves. In the next two case studies, we explore advanced features.

@@

we have a total of 9104 observations and 30 variables, two of which are SUPPORT day 3 physiology score (sps) and 	APACHE III day 3 physiology score (aps). These scores were developed for a period of 180 days, which makes our comparison over 5.56 years exploratory. Therefore, we will assess these two models before using either as our baseline, by comparing them to a full model using all variables. Also, our single imputation introduces bias in the data. As the goal here is purely demonstrative, this bias will be consistent with any of the following models. As such, their comparison is still valid, though any interpretation requires an awareness of said bias. <!--The SUPPORT dataset will be used to demonstrate regularization within casebase, while comparing their absolute risk predictions for a new individual.--> A description of used variables can be found in Table \ref{tab:support1} and a breakdown of each categorical variable in Table \ref{tab:support2}. The original data was obtained from the Department of Biostatistics at Vanderbilt University [@harrell_2020].

@@

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison, beyond a Kaplan-meier curve. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years. Therefore, We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. Next, we will compare each modeling using the Brier score [@graf1999ass]. The Brier score is able to demonstrate both discrimination and calibration in a single model [@graf1999ass] and provide a comparable metric between semi-parametric and parametric methods in the survival context. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score on overall, it will be used in the following absolute risk comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

@@

completeData$d.time<-completeData$d.time/365

@@

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
names(fullDataCox)[names(fullDataCox) == 'V1'] <- 'time'
names(fullDataCox)[names(fullDataCox) == 'V2'] <- 'event'

#here we specify train and test size
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

train <- fullDataCox[train_index, ]
test <- fullDataCox[test_index, ]
coxSPS <- survival::coxph(Surv(time =time , event = event) ~ sps, data = train, x = TRUE)
abcoxSPS <- survival::survfit(coxSPS, newdata = test)

coxAPS <- survival::coxph(Surv(time =time , event = event) ~ aps, data = train, x = TRUE)
abcoxAPS <- survival::survfit(coxAPS, newdata = test)
coxfullForumla<-as.formula(paste("Surv(time = time, event = event) ~ age + sexmale + slos + dzgroupCHF + 
    dzgroupCirrhosis + `dzgroupColon Cancer` + dzgroupComa + 
    dzgroupCOPD + `dzgroupLung Cancer` + `dzgroupMOSF w/Malig` + 
    dzclassCancer + dzclassComa + `dzclassCOPD/CHF/Cirrhosis` + 
    num.co + edu + scoma + avtisst + raceasian + raceblack + 
    racehispanic + raceother + racewhite + hday + diabetes + 
    dementia + cano + cayes + meanbp + wblc + hrt + resp + temp + 
    pafi + alb + bili + crea + sod + ph + glucose + bun + urine + 
    adlp + adlsc "))
coxFull <- survival::coxph(coxfullForumla, data = train, x = TRUE)
abcoxFull <- survival::survfit(coxFull, newdata = test)
# library(prodlim)

@@

colnames(baseCoxComparisonData) <- c("Time", "SPS", "APS","Full")
```

```{r baselines, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:baselines} Absolute risk curves estimated from hazard models. Red: APS only hazard was used to generate the absolute risk. Green: All covariates aside from APS and SPS. Blue: SPS only. Purple: unadjusted Kaplan-Meier curve.",echo=FALSE}
ggplot(as.data.frame(baseCoxComparisonData), aes(Time)) +
  geom_line(aes(y = APS, colour = "APS"), lwd = 1) +
  geom_line(aes(y = Full, col = "Full"), lwd = 1) +
  geom_line(aes(y = SPS, colour = "SPS"), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)",color = "Models")+
   scale_color_manual(values=q4)
```

@@

coxFull$coefficients[which(is.na(coxFull$coefficients))]=0
brierFull <- Score(list("PredictionModel" = coxFull), data = test, formula = Surv(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

brierPlotter<-rbind(
      data.frame(brierScore = brierAPS$Brier$score$Brier[brierAPS$Brier$score$model == "PredictionModel"],times=times,model="APS"),
      data.frame(brierScore = brierFull$Brier$score$Brier[brierFull$Brier$score$model == "PredictionModel"],times=times,model="Full"),
            data.frame(brierScore = brierSPS$Brier$score$Brier[brierSPS$Brier$score$model == "PredictionModel"],times=times,model="SPS")
      )

ggplot(data = brierPlotter, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Survival time (Years)") +
  ylab("Brier-Score")+
   labs(color = "Models")+
  ggtitle("Brier scores over time")+
   scale_color_manual(values=q4[1:3])
```

To get the absolute risk following regularized cox regression, we had to work around limitations in existing packages. To make use of the \pkg{survival} package, we first get our hazard coefficients using \pkg{glmnet}'s coxnet implementation for lasso penalization [@regpathcox]. Next, we fit a cox model using the survival package with the selected variables only, from our previous fit. Finally, we replaced the coefficients in the survival object with those from our glmnet fit, allowing us to make use of the absolute risk tools available in the \pkg{survival} package.

```{r coxNetCode,eval=FALSE,echo=TRUE}
# hazard for cox using glmnet
coxNet <- glmnet::cv.glmnet(x = x, y = y, family = "cox", alpha = 1, standardize = TRUE)
#getting all non-zero covariates
nonzero_covariate_cox <- predict(coxNet, type = "nonzero", s = "lambda.1se")
#keeping coefficients for later
nonzero_coef_cox <- coef(coxNet, s = "lambda.1se")
# cutting dataset to only include nonzero covariates
cleanCoxData <- as.data.frame(cbind(y, x[, nonzero_covariate_cox]))
#create a skeleton object
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData,x=TRUE)
#replace coefficients with regularized ones
coxNet$coefficients <- nonzero_coef_cox@x
```

For our penalized logistic regression model, we opted for the natural log of time, which assumes a Weibull distribution for our event. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where y contains the time and event variables and x contains all other variables we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function.

```{r casebaseDisplayCode, eval=FALSE,echo=TRUE}
#regularized logistic regression to estimate hazard
cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet",
                      time = "d.time", event = "death",
                      formula_time = ~ log(d.time), alpha = 1,
                      ratio = 100, standardize = TRUE, penalty.factor=c(0,rep(1,ncol(x))))

# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull, time = times, newdata = newData, s = "lambda.1se", method = c("numerical"))

```

@@

#remove apsSPS
train<-train[,-c(46,47)]
test<-test[,-c(46,47)]
# x and y will be used to fit the casebase model under glmnet.
y <- data.matrix(train[, c(2, 1)])
colnames(y) <- c("death", "d.time")
x <- data.matrix(train[, -c(2, 1)])
newData <- data.matrix(test[, -c(2, 1)])
```

@@

# fitting a cox model using regular estimation, however we will not keep it.
# this is used more as an object place holder.
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData,x=TRUE)

@@

brierCoxNet <- Score(list("PredictionModel" = coxNet), data = as.data.frame(cbind(test[,c(1,2)],newData[, nonzero_covariate_cox$X1])), formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

@@

```{r cbHazard}, echo = TRUE, eval = eval_cs3}

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 10, standardize = TRUE,penalty.factor=c(0,rep(1,ncol(x))))
```

```{r CBabsolute, echo = TRUE, eval = eval_cs3}
# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull, time = times, newdata = newData, s = "lambda.1se", method = c("numerical"))
```

@@

```{r coefplots,echo=FALSE, eval=eval_cs3, fig.cap="\\label{fig:cs3lolliPlot} A lollipop plot demonstrating the estimated coefficients between the following three models: Cox (green), penalized Cox (red), penalized logisitic (blue). If a coefficient is estimated to be 0, it is removed from the plot"}

@@

lolliplotDots <- rbind(estimatescoxFull, estimatescoxNet,estimatescb)
lolliplotDots$conf.low <- lolliplotDots$estimate
lolliplotDots$conf.high <- lolliplotDots$estimate
lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0
lolliplotDots$term=gsub( '\`', "", as.character(lolliplotDots$term))

lolliplotDots[which(lolliplotDots$estimate==0),c(2,4,5)]<-NA
unChosen<- unique(lolliplotDots[which(is.na(lolliplotDots$estimate)),1])

lolliplotDots[which(lolliplotDots$term %in% unChosen$term),c(2,4,5)]<-NA

lolliplotDots<-lolliplotDots[-which(is.na(lolliplotDots$estimate)),]
dwplot(lolliplotDots)+
  theme(axis.text.y = element_text(size = 11, angle = 0, hjust = 1, vjust = 0),
        strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.855,0.15),legend.text=element_text(size=11)
      )+
  ggtitle("Coefficient plot")+
   labs(color = "Models")+
   scale_color_manual(values=q4[c(2,1,3)])
```

```{r include=TRUE,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3Final} Four absolute risk curves, integrated from the following hazard models: Cox (green), penalized Cox (red), penalized logistic (blue) and unadjusted (Kaplan-Meier, purple)."}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 1) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)",color = "Models")+
  ggtitle("Absolute risk curves")+
   scale_color_manual(values=q4[c(2,1,3,4)])
```

@@

class(cbFull) <- c("singleEventCB",class(cbFull))
brierFinalResults <- Score(list("Cox" = coxFull,"Penalized Cox"=coxNet,"Penalized Logistic"=cbFull), data = test, formula = Hist(time, event != 0) ~ 1, summary = NULL, se.fit = 0L, metrics = "brier", contrasts = FALSE, times =times)

levels(brierFinalResults$Brier$score$model)[levels(brierFinalResults$Brier$score$model)=="Null model"] <- "Unadjusted"
b=brierFinalResults

brierFinalResults$Brier$score$model<- factor(brierFinalResults$Brier$score$model, levels = c("Cox","Penalized Cox","Penalized Logistic", "Unadjusted"))

ggplot(data = brierFinalResults$Brier$score, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Survival time (Years)") +
  ylab("Brier-Score")+
  ggtitle("Brier scores over time")+
  labs(color = "Models")+
  scale_color_manual(values=q4[c(2,1,3,4)])

```

Our first comparison plot examines the covariates that were selected between all three hazard functions. Note that Cox regression without penalization selected more covariates than the other two, but only the intersection of cofficients with values greater than 0 for all models are displayed in \ref{fig:cs3lolliPlot}. A visible tendency is for Cox regression to have larger estimated coefficients. This is expected, as there are no restrictions on the size of coefficients. The smallest coefficient estimates are found in penalized logistic model, aside from cancer state (cano). These values do not inform us of model performance. To do so, we take our hazard functions and use them to predict the absolute risk of each individual in our test set, over time. The probabilites over time for each individual are averaged, resulting in the absolute risk curves in figure \ref{fig:cs3Final}. There is separation between the curves, but it is minimal.  To better demonstrate the differences between the models, we compare brier scores over time. \ref{fig:cs3FinalBrier} the Brier score is lower in our penalized logistic regression model, compared to the other models. This difference, however, is minimal. We can say with relative certainty that any of the hazard models outperform an unadjusted curve, while penalized logistic regression performs the best on this testing dataset.

>>>>

In the first two case studies, we described the basic functionalities of the \pkg{casebase} package: creating population-time plots, fitting parametric models for hazard functions, and estimating the corresponding cumulative incidence curves. In the next case study, we explore an advanced feature.

@@

we have a total of 9104 observations and 30 variables, two of which are SUPPORT day 3 physiology score (sps) and APACHE III day 3 physiology score (aps). These scores were developed for a period of 180 days, which makes our comparison over 5.56 years exploratory. Therefore, we will assess these two models before using either as our baseline, by comparing them to a full model using all variables. Also, our single imputation introduces bias in the data. As the goal here is purely demonstrative, this bias will be consistent with any of the following models. As such, their comparison is still valid, though any interpretation requires an awareness of said bias. <!--The SUPPORT dataset will be used to demonstrate regularization within casebase, while comparing their absolute risk predictions for a new individual.--> A description of covariates and response variables can be found in Table \ref{tab:support1} and a breakdown of each categorical variable in Table \ref{tab:support2}. The original data was obtained from the Department of Biostatistics at Vanderbilt University [@harrell_2020].

@@

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years. Therefore, We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. Next, we will compare each modeling using the Brier score [@graf1999ass]. The Brier score can demonstrate both discrimination and calibration in a single model [@graf1999ass] and provide a comparable metric between semi-parametric and parametric methods in the survival context. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score on overall, it will be used in the following absolute risk comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

@@

completeData$d.time <- completeData$d.time / 365

@@

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
names(fullDataCox)[names(fullDataCox) == "V1"] <- "time"
names(fullDataCox)[names(fullDataCox) == "V2"] <- "event"

# here we specify train and test size
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

train <- fullDataCox[train_index, ]
test <- fullDataCox[test_index, ]
coxSPS <- survival::coxph(Surv(time = time, event = event) ~ sps, data = train, x = TRUE)
abcoxSPS <- survival::survfit(coxSPS, newdata = test)

coxAPS <- survival::coxph(Surv(time = time, event = event) ~ aps, data = train, x = TRUE)
abcoxAPS <- survival::survfit(coxAPS, newdata = test)
coxfullForumla <- as.formula(paste("Surv(time = time, event = event) ~ age + sexmale + slos + dzgroupCHF + 
    dzgroupCirrhosis + `dzgroupColon Cancer` + dzgroupComa + 
    dzgroupCOPD + `dzgroupLung Cancer` + `dzgroupMOSF w/Malig` + 
    dzclassCancer + dzclassComa + `dzclassCOPD/CHF/Cirrhosis` + 
    num.co + edu + scoma + avtisst + raceasian + raceblack + 
    racehispanic + raceother + racewhite + hday + diabetes + 
    dementia + cano + cayes + meanbp + wblc + hrt + resp + temp + 
    pafi + alb + bili + crea + sod + ph + glucose + bun + urine + 
    adlp + adlsc "))
coxFull <- survival::coxph(coxfullForumla, data = train, x = TRUE)
abcoxFull <- survival::survfit(coxFull, newdata = test)
# library(prodlim)

@@

colnames(baseCoxComparisonData) <- c("Time", "SPS", "APS", "Full")
```

```{r baselines, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:baselines} Absolute risk curves estimated from hazard models. Red: APS only hazard was used to generate the absolute risk. Green: All covariates aside from APS and SPS. Blue: SPS only. Purple: unadjusted Kaplan-Meier curve.",echo=FALSE}
ggplot(as.data.frame(baseCoxComparisonData), aes(Time)) +
  geom_line(aes(y = APS, colour = "APS"), lwd = 1) +
  geom_line(aes(y = Full, col = "Full"), lwd = 1) +
  geom_line(aes(y = SPS, colour = "SPS"), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)", color = "Models") +
  scale_color_manual(values = q4)
```

@@

coxFull$coefficients[which(is.na(coxFull$coefficients))] <- 0
brierFull <- Score(list("PredictionModel" = coxFull), data = test, formula = Surv(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

brierPlotter <- rbind(
  data.frame(brierScore = brierAPS$Brier$score$Brier[brierAPS$Brier$score$model == "PredictionModel"], times = times, model = "APS"),
  data.frame(brierScore = brierFull$Brier$score$Brier[brierFull$Brier$score$model == "PredictionModel"], times = times, model = "Full"),
  data.frame(brierScore = brierSPS$Brier$score$Brier[brierSPS$Brier$score$model == "PredictionModel"], times = times, model = "SPS")
)

ggplot(data = brierPlotter, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Survival time (Years)") +
  ylab("Brier-Score") +
  labs(color = "Models") +
  ggtitle("Brier scores over time") +
  scale_color_manual(values = q4[1:3])
```

To get the absolute risk following regularized cox regression, we had to work around limitations in existing packages. To make use of the \pkg{survival} package, we first estimate the hazard coefficients using thecoxnet implementation for lasso penalization in \pkg{glmnet} [@regpathcox]. Next, we fit a cox model using the survival package with the selected variables only, from our regularized fit. Finally, we replaced the coefficients in the survival object with those from our glmnet fit, allowing us to make use of the absolute risk tools available in the \pkg{survival} package.

```{r coxNetCode,eval=FALSE,echo=TRUE}
# hazard for cox using glmnet
coxNet <- glmnet::cv.glmnet(
  x = x, y = y, family = "cox", alpha = 1,
  standardize = TRUE
)
# getting all non-zero covariates
nonzero_covariate_cox <- predict(coxNet,
  type = "nonzero",
  s = "lambda.1se"
)
# keeping coefficients for later
nonzero_coef_cox <- coef(coxNet, s = "lambda.1se")
# cutting dataset to only include nonzero covariates
cleanCoxData <- as.data.frame(cbind(y, x[, nonzero_covariate_cox]))
# create a skeleton object
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ .,
  data = cleanCoxData, x = TRUE
)
# replace coefficients with regularized ones
coxNet$coefficients <- nonzero_coef_cox@x
```

For our penalized logistic regression model, we opted for the natural log of time, which assumes a Weibull distribution for our event. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where *y* contains the time and event variables and *x* contains all covariates we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function.

```{r casebaseDisplayCode, eval=FALSE,echo=TRUE}
# regularized logistic regression to estimate hazard
cbFull <- casebase::fitSmoothHazard.fit(x, y,
  family = "glmnet",
  time = "d.time", event = "death",
  formula_time = ~ log(d.time), alpha = 1,
  ratio = 100, standardize = TRUE,
  penalty.factor = c(0, rep(1, ncol(x)))
)

# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull,
  time = times,
  newdata = newData,
  s = "lambda.1se",
  method = c("numerical")
)
```

@@

# remove apsSPS
train <- train[, -c(46, 47)]
test <- test[, -c(46, 47)]
# x and y will be used to fit the casebase model under glmnet.
y <- data.matrix(train[, c(2, 1)])
colnames(y) <- c("death", "d.time")
x <- data.matrix(train[, -c(2, 1)])
newData <- data.matrix(test[, -c(2, 1)])
```

@@

# fitting a cox model using regular estimation, however we will not keep it.
# this is used more as an object place holder.
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData, x = TRUE)

@@

brierCoxNet <- Score(list("PredictionModel" = coxNet), data = as.data.frame(cbind(test[, c(1, 2)], newData[, nonzero_covariate_cox$X1])), formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

@@

```{r cbHazard}, echo = FALSE, eval = eval_cs3}

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 10, standardize = TRUE, penalty.factor = c(0, rep(1, ncol(x))))
```

```{r CBabsolute, echo = FALSE, eval = eval_cs3}
# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull, time = times, newdata = newData, s = "lambda.1se", method = c("numerical"))
```

@@

```{r coefplots,echo=FALSE, eval=eval_cs3, fig.cap="\\label{fig:cs3lolliPlot} A lollipop plot demonstrating the estimated coefficients between the following three models: Cox (green), penalized Cox (red), penalized logisitic (blue). If a coefficient is estimated to be 0, in any model, it is removed from the plot for simplicity."}

@@

lolliplotDots <- rbind(estimatescoxFull, estimatescoxNet, estimatescb)
lolliplotDots$conf.low <- lolliplotDots$estimate
lolliplotDots$conf.high <- lolliplotDots$estimate
lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0
lolliplotDots$term <- gsub("\`", "", as.character(lolliplotDots$term))

lolliplotDots[which(lolliplotDots$estimate == 0), c(2, 4, 5)] <- NA
unChosen <- unique(lolliplotDots[which(is.na(lolliplotDots$estimate)), 1])

lolliplotDots[which(lolliplotDots$term %in% unChosen$term), c(2, 4, 5)] <- NA

lolliplotDots <- lolliplotDots[-which(is.na(lolliplotDots$estimate)), ]
dwplot(lolliplotDots) +
  theme(
    axis.text.y = element_text(size = 11, angle = 0, hjust = 1, vjust = 0),
    strip.text.x = element_blank(),
    strip.background = element_rect(colour = "white", fill = "white"),
    legend.position = c(.855, 0.15), legend.text = element_text(size = 11)
  ) +
  ggtitle("Coefficient plot") +
  labs(color = "Models") +
  scale_color_manual(values = q4[c(2, 1, 3)])
```

```{r include=TRUE,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3Final} Four absolute risk curves, integrated from the following hazard models: Cox (green), penalized Cox (red), penalized logistic (blue) and unadjusted (Kaplan-Meier, purple)."}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 1) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 1) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 1) +
  labs(y = "Probability of Death", x = "Survival time (Years)", color = "Models") +
  ggtitle("Absolute risk curves") +
  scale_color_manual(values = q4[c(2, 1, 3, 4)])
```

@@

class(cbFull) <- c("singleEventCB", class(cbFull))
brierFinalResults <- Score(list("Cox" = coxFull, "Penalized Cox" = coxNet, "Penalized Logistic" = cbFull), data = test, formula = Hist(time, event != 0) ~ 1, summary = NULL, se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

levels(brierFinalResults$Brier$score$model)[levels(brierFinalResults$Brier$score$model) == "Null model"] <- "Unadjusted"
b <- brierFinalResults

brierFinalResults$Brier$score$model <- factor(brierFinalResults$Brier$score$model, levels = c("Cox", "Penalized Cox", "Penalized Logistic", "Unadjusted"))

ggplot(data = brierFinalResults$Brier$score, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Survival time (Years)") +
  ylab("Brier-Score") +
  ggtitle("Brier scores over time") +
  labs(color = "Models") +
  scale_color_manual(values = q4[c(2, 1, 3, 4)])
```

Our first comparison plot examines the covariates that were selected between all three hazard functions. Note that Cox regression without penalization selected more covariates than the other two, but only the intersection of coefficients with values greater than 0 are displayed in \ref{fig:cs3lolliPlot}. A visible tendency is for Cox regression to have larger estimated coefficients. This is expected, as there are no restrictions on the size of coefficients. The smallest coefficient estimates are found in penalized logistic model, aside from cancer state (cano). These values do not inform us of model performance. To do so, we take our hazard functions and use them to predict the absolute risk of everyone in our test set, over time. The probabilities over time for each individual are averaged, resulting in the absolute risk curves in figure \ref{fig:cs3Final}. There is separation between the curves, but it is minimal. To better demonstrate the differences between the models, we compare brier scores over time. \ref{fig:cs3FinalBrier} The Brier score is lower in our penalized logistic regression model, compared to the other models. This difference, however, is minimal. We can say with relative certainty that any of the hazard models outperform an unadjusted curve, while penalized logistic regression performs the best on this testing dataset.
